\LARGE Untangling the link between traits, size and growth rate in plants

### 2016-07-28 ms/ms.tex
Remove data component

% allows for numbered referencing, as refquired by ecol letters
\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers} % comment to remove numbers in reference list

@@

\author[1]{Daniel S. Falster}
\author[1]{Richard G. FitzJohn}
\author[2]{Joe Wright}

\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
\affil[2]{{\footnotesize Center for Tropical Forest Science, Smithsonian Tropical Research Institute, Panama, Republic of Panama}}
\renewcommand\Authands{ and }
\date{\vspace{-3em}}

@@

Ulitmately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{sillett-2010, king-2011} or diameter growth \citep{herault-2011}.  In contrast, basal-area growth continues to increase with size \citep{sillett-2010, stephenson-2014}, due to an increasing influence of stem turnover. All growth measures decrease sharply with size when expressed as relative growth rates \citep{rees-2010, iida-2014}. Shade tolerance tends to decrease with increasing size \citep{lusk-2006}, due to increasing costs of mainatining stems \citep{givnish-1988}

@@

The results of this new analysis
(Fig. \ref{f-BCI}) align closely with our theoretical predictions  showing how the effect of traits on growth
changes with plant size, with different traits exhibiting distinct
size-dependent effects.

\subsection{Empirical motivation}\label{fresh-empirical-motivation}

To help illustrate the empirical patterns, we begin with a fresh analysis of
growth dynamics in the long-term plot at Barro Colorado Island in Panama,
where the growth rates for over 100 species have been recorded over the last
30 years. Data from a long term plot \citep{condit-2012} are combined
with results of experiment on seedlings \citep{kitajima-2013}, to better understand
size-related patterns of growth. A prior analysis by
 \citet{wright-2010} found a moderate effect of wood density on
growth rate, but almost no effect of LMA, and only a weak effect of HMAX. More recently,
 \citet{ruger-2012} showed

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Effect of lma stronger than previously thought,
\item
  Bridges from seedling studies to adults
\end{itemize}

Results stronger than previously reported.

Predictions
are compared against data from both a long-term
plot \citep{condit-2012} and field
experiment \citep{kitajima-2013} at Barro Colorado Island, Panama.
We use a new method to estimate species potential growth rate,
accounting for size specific effects (Fig. \ref{fS-quantile_examples},
see Methods for details). This new method allowed us to extract a more
definite signal of trait effects than in previous
studies \citep{wright-2010}.

@@

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size (Figs. \ref{f-BCI}, \ref{f-lma_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

@@

 Greater height at maturation ($h_m$)
can thus lead to a growth advantage by increasing
$\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
among species, because most species are focusing on growth (Fig.
\ref{f-BCI}). At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate (Fig. \ref{f-BCI}).

@@

\begin{figure}[ht]
\centering
\includegraphics{BCI_data.pdf}
\caption{\textbf{The relationship between traits and potential growth
rate varies with plant size.} For XXX species from tropical rain forest
in Panama, we estimated the potential growth rate of individual's in
each species at a series of diameters $D$, indicated along right hand
side. The size of circles in each panel indicates the number of data
points used to estimate potential growth rate. Traits values were
calculated from representative individuals in each species.
\label{f-BCI}}
\end{figure}

@@

\begin{figure}[ht]
\centering
\includegraphics{SI_quantile_examples.pdf}
\caption{\textbf{Estimating potential growth rate via quantile
regression.} Example for 4 species. Growth rate estimated at six sizes,
by fitting lines within given size intervals. Lines fitted to 99th
quantile, but only when n \textgreater{} 500. Except for left-hand
panel, estimated points in centre of interval.
\label{fS-quantile_examples}}
\end{figure}

>>>>

% allows for numbered referencing, as required by ecol letters
\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers} % comment to remove numbers in reference list

@@

\author[1]{Daniel S. Falster}
\author[1]{Richard G. FitzJohn}

\affil[1]{{\footnotesize Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}}
\renewcommand\Authands{ and }
\date{\vspace{-3em}}

@@

Ultimately, trait based understanding integrated wiht model of plant growth, also able to cpature patterns of growth through ontogeny. Growth rates tend to show hump-shaped relationship with size, when expressed as either height \citep{sillett-2010, king-2011} or diameter growth \citep{herault-2011}.  In contrast, basal-area growth continues to increase with size \citep{sillett-2010, stephenson-2014}, due to an increasing influence of stem turnover. All growth measures decrease sharply with size when expressed as relative growth rates \citep{rees-2010, iida-2014}. Shade tolerance tends to decrease with increasing size \citep{lusk-2006}, due to increasing costs of mainatining stems \citep{givnish-1988}

@@

@@

Unlike eq. \ref{eq:RGR}, eqs. \ref{eq:dhdt}, \ref{eq:dast} and
\ref{eq:dDdt} predict a relationship between $\phi$ and growth that
changes with plant size ( \ref{f-lma_growth_size}).
Decreasing $\phi$ has two impacts on growth rate. First, lower
$\phi$ increases marginal leaf deployment
($\textrm{d}a_\textrm{l} / \textrm{d}m_\textrm{a}$ by economising on
construction costs. Second, lower $\phi$ decreases net production
($\textrm{d}P / \textrm{d}t$), due to increased leaf turnover. Whether
lower $\phi$ increases growth thus depends on the relative magnitude
of these two effects. When plants are small the effect on leaf
deployment rate is bigger and so decreasing $\phi$ increases growth
rate. When plants are large, the influence of $\phi$ on leaf
deployment rate is diminished, because the costs of building other
supportive tissues (other terms in eq. \ref{eq:daldmt}) are larger (Fig.
\ref{f-mass_fraction}). The net result is that at larger sizes, low
$\phi$ is no longer advantageous for growth (Fig.
\ref{f-lma_growth_size}d).

@@

 Greater height at maturation ($h_m$)
can thus lead to a growth advantage by increasing
$\frac{\textrm{d}m_\textrm{a}}{\textrm{d}b}$ . At smaller sizes, there is no differentiation
among species, because most species are focusing on growth. At larger sizes, individuals of some species are
allocating a majority of their surplus energy to growth, leaving a
strong signal in potential growth rate.

@@

@@
